\begin{abstract}
Large Language Models (LLMs) have demonstrated remarkable capabilities in general natural language tasks but frequently fail in formal reasoning domains requiring exact, deterministic inferences—such as multi-step chemical synthesis, Euclidean flat geometry proofs, and algebraic manipulation. In these mathematical environments, LLMs suffer from semantic drift, visual-geometric hallucination, and logical circularity. To resolve these limitations, we present \textbf{Omni-IPS} (\textit{Neuro-Symbolic \& GraphRAG Multi-Domain Intelligent Problem Solver}), a hybrid system that bridges statistical language models and deterministic logical reasoning engines. Omni-IPS leverages a multi-agent parser utilizing a specialized grammar-based balanced parenthesis scanner to extract strict logical propositions from natural language queries. These queries are routed via a Neuro-Symbolic RAG mechanism using a dual-database storage: high-dimensional vector representations in Qdrant for semantic mapping, and a rigid ontological schema in Neo4j representing canonical facts and rules. Once aligned, a domain-agnostic Forward and Backward Chaining inference engine executes step-by-step reasoning in a closed-world assumption environment, producing completely verified, hallucination-free proof traces. Finally, an online explainability agent translates the dry logical trace into a rich, structured educational response streamed in real-time to the user interface. We demonstrate that Omni-IPS achieves 100\% correctness in symbolic proof verification across high-school level problems in chemistry, geometry, and algebra, significantly outperforming pure LLM prompting baselines.
\end{abstract}
