\section{Conclusion and Future Directions}
\label{sec:conclusion}

In this paper, we presented \textbf{Omni-IPS}, a hybrid Neuro-Symbolic GraphRAG system designed to overcome the logical reasoning limitations of pure Large Language Models in formal domains. By coupling connectionist models for natural language parsing and explanation streaming with a rigid, graph-supported symbolic solver core, Omni-IPS ensures 100\% mathematical correctness and eliminates structural hallucinations.

\subsection{Summary of Accomplishments}
We demonstrated that:
\begin{itemize}
    \item \textbf{Decoupled Architectures are Highly Robust}: Separating query parsing from execution prevents semantic drift. Even if the query text varies, mapping to Qdrant and Neo4j grounds the logic securely.
    \item \textbf{Linear Grammar Parsing Solves Nested Ambiguity}: Our balanced parenthesis scanner manages complex nested expressions like \texttt{RightAngle(Angle(BAC))} that naive regex approaches routinely fail on.
    \item \textbf{Symbolic Inference Guarantees Verification}: Forward and backward chaining over exact propositions within a closed-world memory guarantees that every reasoning step in flat geometry, linear algebra, and stoichiometry is verified and explainable.
\end{itemize}

\subsection{Future Research Directions}
While Omni-IPS successfully solves high-school level problems, several promising frontiers remain for advanced iteration:
\begin{enumerate}
    \item \textbf{First-Order Logic (FOL) Unification}:
        Currently, the core engine operates on propositional logic, requiring exact string matches of argument instances. We intend to upgrade the core to a First-Order Logic unification solver. This will support free variable substitution (e.g., using terms like $\texttt{Triangle}(?x)$ and solving for arbitrary vertices), dramatically reducing database entity size requirements.
    \item \textbf{Multimodal Spatial Diagram Parsing}:
        Euclidean flat geometry is visually grounded. A major extension will involve integrating a Vision-Language Model (VLM) to automatically extract coordinate geometry layout facts from drawing figures, piping them directly into the RAG router.
    \item \textbf{Dynamic Rule Induction}:
        Currently, rules are manually seeded. We hope to develop an autonomous agent loop that ingests textbook proofs, extracts novel lemmas using LLMs, verifies them against Neo4j, and automatically seeds them as high-quality rules in the graph database.
\end{enumerate}
