\section{Symbolic Reasoning Engine Core}
\label{sec:reasoning_engine}

The computational core of Omni-IPS is its domain-agnostic, deterministic symbolic reasoning engine in \texttt{core\_engine/solver.py}. By operating strictly on concrete propositions in a closed-world assumption (CWA), the engine avoids semantic drift, math errors, and reasoning loops~\cite{raedt2020neuro}. We implement two core reasoning strategies: \textbf{Forward Chaining} (data-driven reasoning) and \textbf{Backward Chaining} (goal-driven hypothesis proving), grounding our logic in classic algorithmic search paradigms~\cite{russell2020artificial}.

\subsection{Mathematical Formulations}
Let $\Facts$ be the universe of all valid logical facts in a given domain, and let $\Rules$ be the set of rules loaded from the database. Each rule $R \in \Rules$ is a tuple defined as:
\begin{equation}
    R = \langle \text{id}, \text{name}, \mathcal{I}_R, \mathcal{O}_R \rangle
\end{equation}
where $\mathcal{I}_R \subset \Facts$ is the set of antecedents (prerequisites) and $\mathcal{O}_R \subset \Facts$ is the set of consequents (conclusions). Let $\WM_k \subseteq \Facts$ be the state of the Working Memory at step $k$, and let $\mathcal{A}_k$ be the set of fired rule IDs up to step $k$.

\subsubsection{Forward Chaining State Transitions}
Forward chaining starts with initial facts $\WM_0$ and applies rules iteratively until a saturation state is reached or the goal is satisfied.
Let $G \in \Facts$ be the target goal fact. The initialization states are:
\begin{align}
    \WM_0 &= \text{initial\_facts} \\
    \mathcal{A}_0 &= \emptyset
\end{align}

At each iteration $k \ge 0$, the system searches for a rule $R \in \Rules$ that satisfies two conditions:
\begin{align}
    \mathcal{I}_R &\subseteq \WM_k \\
    R.\text{id} &\notin \mathcal{A}_k
\end{align}
If such a rule is matched, the state updates are:
\begin{align}
    \WM_{k+1} &= \WM_k \cup \mathcal{O}_R \\
    \mathcal{A}_{k+1} &= \mathcal{A}_k \cup \{ R.\text{id} \}
\end{align}
The process terminates under two conditions:
\begin{itemize}
    \item \textbf{Goal Satisfied}: $G \in \WM_{k+1}$. The search terminates immediately, returning success.
    \item \textbf{Saturation}: $\forall R \in \Rules$, either $\mathcal{I}_R \not\subseteq \WM_k$ or $R.\text{id} \in \mathcal{A}_k$. No new facts can be inferred ($\WM_{k+1} = \WM_k$), returning failure.
\end{itemize}

\subsubsection{Backward Chaining Recursive Proving}
Backward Chaining operates in reverse, starting with a goal $G$ and recursively identifying subgoals. Let $\mathcal{V}$ be the set of active subgoals in the recursion stack, acting as a cycle-detection guard to prevent infinite loops. The proof of a subgoal $g \in \Facts$ is recursively defined as:
\begin{equation}
    \text{prove}(g) = (g \in \WM_k) \lor \left( g \notin \mathcal{V} \land \bigvee_{R \in \text{Candidates}(g)} \left( \bigwedge_{a \in \mathcal{I}_R} \text{prove}(a) \right) \right)
\end{equation}
where $\text{Candidates}(g) = \{ R \in \Rules \mid g \in \mathcal{O}_R \}$ is the set of rules capable of producing the subgoal $g$. If a candidate rule $R$ successfully proves all of its antecedents $a \in \mathcal{I}_R$, it fires, adding its outcomes to working memory:
\begin{equation}
    \WM_{k+1} = \WM_k \cup \mathcal{O}_R
\end{equation}

\subsection{Reasoning Algorithms}
The complete logic for Forward Chaining and Backward Chaining is detailed in Algorithm~\ref{alg:forward_chaining} and Algorithm~\ref{alg:backward_chaining} respectively.

\begin{algorithm}[h]
\caption{Forward Chaining Inference Engine}
\label{alg:forward_chaining}
\begin{algorithmic}[1]
\Procedure{SolveForward}{$\text{initial\_facts}, \text{rules}, G$}
    \State $\WM \gets \text{set}(\text{initial\_facts})$
    \State $\mathcal{A} \gets []$, $\mathcal{T} \gets []$
    \If{$G \neq \emptyset$ \textbf{and} $G \in \WM$}
        \State \textbf{return} $\text{InferenceResult}(\text{goal\_reached}=\text{True}, \WM, \mathcal{T}, \mathcal{A})$
    \EndIf
    \State $new\_facts \gets \text{True}$
    \While{$new\_facts$}
        \State $new\_facts \gets \text{False}$
        \For{$R$ \textbf{in} $rules$}
            \If{$R.\text{id} \notin \mathcal{A}$ \textbf{and} $\mathcal{I}_R \subseteq \WM$}
                \State $inferred \gets \{ f \in \mathcal{O}_R \mid f \notin \WM \}$
                \State $\WM \gets \WM \cup \mathcal{O}_R$
                \State $\text{Append}(\mathcal{A}, R.\text{id})$
                \State $\text{Append}(\mathcal{T}, \text{Step}(R.\text{id}, inferred))$
                \State $new\_facts \gets \text{True}$
                \If{$G \neq \emptyset$ \textbf{and} $G \in \WM$}
                    \State \textbf{return} $\text{InferenceResult}(\text{goal\_reached}=\text{True}, \WM, \mathcal{T}, \mathcal{A})$
                \EndIf
            \EndIf
        \EndFor
    \EndWhile
    \State \textbf{return} $\text{InferenceResult}(\text{goal\_reached}=\text{False}, \WM, \mathcal{T}, \mathcal{A})$
\EndProcedure
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[h]
\caption{Backward Chaining Inference Engine}
\label{alg:backward_chaining}
\begin{algorithmic}[1]
\Procedure{SolveBackward}{$\text{initial\_facts}, \text{rules}, G$}
    \State $\WM \gets \text{set}(\text{initial\_facts})$
    \State $\mathcal{A} \gets []$, $\mathcal{T} \gets []$, $\mathcal{V} \gets \text{set}()$
    \Function{Prove}{$g$}
        \If{$g \in \WM$} \State \textbf{return} \text{True} \EndIf
        \If{$g \in \mathcal{V}$} \State \textbf{return} \text{False} \EndIf
        \State $\mathcal{V} \gets \mathcal{V} \cup \{ g \}$
        \State $candidates \gets \{ R \in \text{rules} \mid g \in \mathcal{O}_R \}$
        \For{$R$ \textbf{in} $candidates$}
            \State $proven \gets \text{True}$
            \For{$a$ \textbf{in} $\mathcal{I}_R$}
                \If{\textbf{not} $\text{Prove}(a)$}
                    \State $proven \gets \text{False}$
                    \State \textbf{break}
                \EndIf
            \EndFor
            \If{$proven$}
                \State $inferred \gets \{ f \in \mathcal{O}_R \mid f \notin \WM \}$
                \State $\WM \gets \WM \cup \mathcal{O}_R$
                \If{$R.\text{id} \notin \mathcal{A}$}
                    \State $\text{Append}(\mathcal{A}, R.\text{id})$
                    \State $\text{Append}(\mathcal{T}, \text{Step}(R.\text{id}, inferred))$
                \EndIf
                \State $\mathcal{V} \gets \mathcal{V} \setminus \{ g \}$
                \State \textbf{return} \text{True}
            \EndIf
        \EndFor
        \State $\mathcal{V} \gets \mathcal{V} \setminus \{ g \}$
        \State \textbf{return} \text{False}
    \EndFunction
    \State $goal\_reached \gets \text{Prove}(G)$
    \State \textbf{return} $\text{InferenceResult}(goal\_reached, \WM, \mathcal{T}, \mathcal{A})$
\EndProcedure
\end{algorithmic}
\end{algorithm}
